# Test script

```{r setup}
library(tidyverse)
library(tidync)
library(gdx)
library(gdxdt)  # From Alois GitHub repo
library(ggnewscale)
library(devtools)
library(quitte)

file.name <- "../results/networks/elec_s_3_ec_lcopt_Ep-1H.nc"

devtools::load_all(file.path("C:/Users/adrianod/Documents/PhD/Modelling/tidync"))

source("readData.R")
source("readData2.R")
source("position_stack_and_nudge.R")

```

# Read in variables

## Hourly data

```{r, warning = F}

# Generation
vars.h.gen <- c("generators_t_p",
                "buses_t_marginal_price",
                "loads_t_p")

data.h.gen <- readData(file.name, vars.h.gen, hourly = T)

# Storage
vars.h.stor <- c("storage_units_t_p")

data.h.stor <- readData(file.name, vars.h.stor, hourly =T)

```

## Yearly data

```{r, warning = F}

vars.a.gen <- c("generators_p_nom_opt",
                "generators_capital_cost",
                "generators_marginal_cost")

data.a.gen <- readData(file.name, vars.a.gen, hourly = F)

```

# Plot some stuff

## Colormap

```{r}

color.tech <- c(
  "solar" = "#ffcc00",
  "onwind" = "#337fff",
  "offwind-ac" = "#63BFE0",
  "offwind-dc" = "#63BFE0",
  "ror" = "#191999",
  "biomass" = "#005900",
  "OCGT" = "#616138",
  "CCGT" = "#999959",
  "oil" = "#663a00",
  "coal" = "#845353",
  "lignite" = "#0c0c0c",
  "nuclear" = "#ff33ff"
)

color.tech.stor <- c(
  "PHS" = "#117FA5",
  "battery" = "#8E5FBE",
  "hydro" = "#7175D8",
  "H2" = "#74E0F2"
)
  
```


## Capacity

```{r}

data.plot <- readData2("generators_p_nom_opt",
                      c("region", "bus", "technology")) %>% 
  order.levels(technology = names(color.tech))
  
p <- ggplot() + 
  geom_bar(data = data.plot,
           mapping = aes(x = bus, y = generators_p_nom_opt, fill = technology),
           stat = "identity",
           position = "stack") + 
  scale_fill_manual(values = color.tech)

print(p)

data.plot <- data.plot %>% 
  group_by(region, technology) %>% 
  summarise(generators_p_nom_opt = sum(generators_p_nom_opt))

p <- ggplot() + 
  geom_bar(data = data.plot,
           mapping = aes(x = region, y = generators_p_nom_opt, fill = technology),
           stat = "identity",
           position = "stack") + 
  scale_fill_manual(values = color.tech)

print(p)

```

## Generation

```{r}

data.plot <- readData2("generators_t_p",
                      c("region", "bus", "technology")) %>% 
  group_by(region, bus, technology) %>% 
  summarise(generators_t_p = sum(generators_t_p)) %>% 
  order.levels(technology = names(color.tech))
  
p <- ggplot() + 
  geom_bar(data = data.plot,
           mapping = aes(x = bus, y = generators_t_p, fill = technology),
           stat = "identity",
           position = "stack") + 
  scale_fill_manual(values = color.tech)

print(p)

data.plot <- data.plot %>% 
  group_by(region, technology) %>% 
  summarise(generators_t_p = sum(generators_t_p))

p <- ggplot() + 
  geom_bar(data = data.plot,
           mapping = aes(x = region, y = generators_t_p, fill = technology),
           stat = "identity",
           position = "stack") + 
  scale_fill_manual(values = color.tech)

print(p)

```

## Load and power (one week in January)

```{r}

t.range <- seq(168,2*168)  # second week

data.plot.load <- readData2("loads_t_p",
                            c("region", "bus")) %>% 
  filter(snapshots %in% t.range)

data.plot.power <- readData2("generators_t_p",
                             c("region", "bus", "technology")) %>% 
  filter(snapshots %in% t.range) %>% 
  order.levels(technology = names(color.tech))

p <- ggplot() +
  geom_area(data = data.plot.power,
            mapping = aes(x = snapshots, y = generators_t_p, fill = technology)) + 
  geom_line(data = data.plot.load,
            mapping = aes(x = snapshots, y = loads_t_p)) +
  scale_fill_manual(values = color.tech) + 
  facet_wrap(~bus)
print(p)

data.plot.load <- data.plot.load %>% 
  group_by(region, snapshots) %>% 
  summarise(loads_t_p = sum(loads_t_p))

data.plot.power <- data.plot.power %>% 
  group_by(region, technology, snapshots) %>% 
  summarise(generators_t_p = sum(generators_t_p))

p <- ggplot() +
  geom_area(data = data.plot.power,
            mapping = aes(x = snapshots, y = generators_t_p, fill = technology)) + 
  geom_line(data = data.plot.load,
            mapping = aes(x = snapshots, y = loads_t_p)) + 
  scale_fill_manual(values = color.tech)
print(p)

# With storage
data.plot.stor <- readData2("storage_units_t_p",
                            c("region", "bus", "technology")) %>% 
  filter(snapshots %in% t.range) %>% 
  group_by(region, technology, snapshots) %>% 
  summarise(storage_units_t_p = sum(storage_units_t_p))

data.plot.genstor <- bind_rows(data.plot.gen %>%
                                 rename(value = generators_t_p),
                               data.plot.stor %>% 
                                 rename(value = storage_units_t_p)) %>% 
  order.levels(technology = names(c(color.tech.stor, color.tech)))

p <- ggplot() +
  geom_area(data = data.plot.genstor,
            mapping = aes(x = snapshots, y = value, fill = technology)) + 
  geom_line(data = data.plot.load,
            mapping = aes(x = snapshots, y = loads_t_p)) + 
  scale_fill_manual(values = c(color.tech.stor, color.tech))
print(p)

```
## Electricity price (one week in January)

```{r}

data.plot.price <- readData2(c("buses_t_marginal_price", "loads_t_p"),
                             c("region", "bus")) %>% 
  filter(snapshots %in% t.range)

p <- ggplot(data = data.plot.price) +
  geom_line(aes(x = snapshots, y = buses_t_marginal_price)) +
  facet_wrap(~bus)

print(p)

data.plot.price.weighted <- data.plot.price %>% 
  group_by(region, snapshots) %>% 
  summarise(price.weighted = sum(buses_t_marginal_price * loads_t_p) / sum(loads_t_p))
  
p <- ggplot(data = data.plot.price.weighted) +
  geom_line(aes(x = snapshots, y = price.weighted))

print(p)

```

## Price duration curve

```{r}

data.plot.price.weighted <- readData2(c("buses_t_marginal_price", "loads_t_p"),
                                      c("region", "bus")) %>% 
  group_by(region, snapshots) %>% 
  summarise(price.weighted = sum(buses_t_marginal_price * loads_t_p) / sum(loads_t_p)) %>% 
  arrange(desc(price.weighted))

p <- ggplot(data = data.plot.price.weighted) + 
  geom_line(aes(x = seq(1,8760), y = price.weighted))

print(p)

```

# Basic economics

## Total payments = Total revenues (w/o investments, follows from power balance)

Show that market price-weighted load (consumer payments) is equal to
market price-weighted generation (generator revenues) + congestion rent!

```{r}

# Consumer payments
data.plot.pay <- readData(file.name,
                          c("loads_t_p", "buses_t_marginal_price"),
                          hourly = T) %>%
  summarise(total.payment = sum(loads_t_p * buses_t_marginal_price))

# Generator revenues
data.plot.rev <- readData(file.name,
                          c("generators_t_p", "buses_t_marginal_price"),
                          hourly = T) %>% 
  summarise(total.revenue = sum(generators_t_p * buses_t_marginal_price))

# Storage revenues
data.plot.stor.rev <- readData(file.name,
                               c("storage_units_t_p", "buses_t_marginal_price"),
                               hourly = T) %>% 
  summarise(total.revenue = sum(storage_units_t_p * buses_t_marginal_price))

# Congestion rent
test1 <- readData2(vars = c("lines_t_p0", "lines_t_p1"),
                   dims = c("lines"))

test2 <- readData2(vars = c("lines_bus0", "lines_bus1"),
                   dims = "lines")

test3 <- full_join(test1, test2) %>% 
  separate(lines_bus0,
           into = c("region0", "bus0"),
           sep = "\\s") %>% 
  separate(lines_bus1,
           into = c("region1", "bus1"),
           sep = "\\s")

price <- readData2(vars = "buses_t_marginal_price",
                   dims = c("region", "bus"))

temp1 <- full_join(test3, price, by = c("region0" = "region",
                                       "bus0" = "bus",
                                       "snapshots")) %>% 
  filter(!is.na(lines_t_p0)) %>% 
  summarise(sum1 = sum(lines_t_p0*buses_t_marginal_price))

temp2 <- full_join(test3, price, by = c("region1" = "region",
                                       "bus1" = "bus",
                                       "snapshots")) %>% 
  filter(!is.na(lines_t_p0)) %>% 
  summarise(sum2 = sum(lines_t_p0*buses_t_marginal_price))

data.plot.cong.rent <- bind_cols(temp1, temp2) %>% 
  transmute(cong.rent = abs(sum1 - sum2))

data.plot <- bind_rows(
  data.plot.rev %>% 
    mutate(variable = "generation revenue") %>% 
    rename(value = total.revenue),
  data.plot.stor.rev %>% 
    mutate(variable = "storage revenue") %>% 
    rename(value = total.revenue),
  data.plot.cong.rent %>% 
    mutate(variable = "congestion rent") %>% 
    rename(value = cong.rent)
)

p <- ggplot() + 
  geom_bar(data = data.plot,
           mapping = aes(x = 0, y = value, fill = variable),
           stat = "identity") +
  scale_fill_discrete(name = "Supply side") +
  new_scale_fill() + 
  geom_bar(data = data.plot.pay,
           mapping = aes(x = 1, y = total.payment, fill = "Total payments"),
           stat = "identity") + 
  scale_fill_manual(name = "Demand side",
                    values = c("Total payments" = "gray")) + 
  ggtitle("Equality of payments and revenues, without costs!")

print(p)

```

## Zero profit conditions

### Generation (LCOE = MV)

```{r}

data.costs <- readData2(c("generators_capital_cost", "generators_marginal_cost"),
                        c("region", "bus", "technology"))

data.gen <- readData2(c("generators_p_nom_opt", "generators_t_p"),
                      c("region", "bus", "technology"))

data.price <- readData2("buses_t_marginal_price",
                        c("region", "buses"))

data.lcoe <- data.gen %>% 
  full_join(data.costs) %>% 
  group_by(technology, bus) %>% 
  summarise(fix = (sum(generators_capital_cost * 1/8760*generators_p_nom_opt)/sum(generators_t_p)),
            var = (sum(generators_marginal_cost * generators_t_p)/sum(generators_t_p))) %>% 
  pivot_longer(cols = c(fix, var), names_to = "lcoe_part") %>% 
  order.levels(lcoe_part = c("var", "fix"))

data.mv <- data.gen %>% 
  full_join(data.price) %>% 
  group_by(technology, bus) %>% 
  summarise(mv = case_when(sum(generators_t_p) > 1E-3 ~ sum(buses_t_marginal_price*generators_t_p)/sum(generators_t_p),
                           TRUE ~ 0))

p <- ggplot() + 
  geom_bar(data = data.lcoe,
           mapping = aes(x = bus, y = value, fill = lcoe_part),
           stat = "identity",
           width = 0.3,
           position = position_stack_and_nudge(x = -0.2)) +
  scale_fill_discrete(name = "Investment side",
                      labels = c("var" = "LCOE (variable)",
                                 "fix" = "LCOE (fixed)")) + 
  new_scale_fill() + 
  geom_bar(data = data.mv %>% filter(technology != "oil"),
           mapping = aes(x = bus, y = mv, fill = "Market value"),
           stat = "identity",
           width = 0.3,
           position = position_stack_and_nudge(x = 0.2)) +
  scale_fill_manual(name = "Revenue side",
                    values = c("Market value" = "grey50")) +
  facet_wrap(~technology) +
  ggtitle("LCOE = Market value, only for extendable technologies")

print(p)

```

### Transmission

```{r}

data.costs <- readData2("lines_capital_cost",
                        "lines")

data.lines <- readData2(c("lines_s_nom_opt", "lines_s_nom"),
                        "lines") %>% 
  mutate(temp = lines_s_nom_opt - lines_s_nom)

test1 <- readData2(vars = c("lines_t_p0", "lines_t_p1"),
                   dims = c("lines"))

test2 <- readData2(vars = c("lines_bus0", "lines_bus1"),
                   dims = "lines")

test3 <- full_join(test1, test2) %>% 
  separate(lines_bus0,
           into = c("region0", "bus0"),
           sep = "\\s") %>% 
  separate(lines_bus1,
           into = c("region1", "bus1"),
           sep = "\\s")

price <- readData2(vars = "buses_t_marginal_price",
                   dims = c("region", "bus"))

temp1 <- full_join(test3, price, by = c("region0" = "region",
                                       "bus0" = "bus",
                                       "snapshots")) %>% 
  filter(!is.na(lines_t_p0)) %>% 
  group_by(lines) %>% 
  summarise(sum1 = sum(lines_t_p0*buses_t_marginal_price))

temp2 <- full_join(test3, price, by = c("region1" = "region",
                                       "bus1" = "bus",
                                       "snapshots")) %>% 
  filter(!is.na(lines_t_p0)) %>% 
  group_by(lines) %>% 
  summarise(sum2 = sum(lines_t_p0*buses_t_marginal_price))

data.plot.cong.rent <- full_join(temp1, temp2) %>% 
  mutate(cong.rent = abs(sum1 - sum2))

temp <- full_join(data.plot.cong.rent, data.costs) %>% 
  full_join(data.lines) %>% 
  mutate(invest = lines_capital_cost * temp) %>% 
  summarise(cong.rent = sum(cong.rent),
            invest = sum(invest))

```

### Storage

## Market value vs. Marginal costs

```{r}

# Actual power
data.p <- data.hourly %>%
  filter(variable == "generators_t_p") %>% 
  rename(generators_t_p = value) %>% 
  select(region, bus, technology, snapshots, generators_t_p)

# Marginal price
data.price <- data.hourly %>% 
  filter(variable == "buses_t_marginal_price") %>% 
  rename(buses_t_marginal_price = value) %>% 
  select(region, bus, snapshots, buses_t_marginal_price)

# Market value
data.mv <- full_join(data.p, data.price) %>% 
  group_by(region, bus, technology) %>% 
  summarise(sum.revenue = sum(generators_t_p * buses_t_marginal_price),
            sum.gen = sum(generators_t_p)) %>%
  group_by(region, bus, technology) %>%
  transmute(value = sum.revenue / sum.gen,
            variable = "market value")

# Capital costs
data.cc <- data.yearly %>% 
  filter(variable %in% c("generators_capital_cost", "generators_p_nom_opt")) %>% 
  select(!unit) %>% 
  pivot_wider(names_from = variable, values_from = value)

data.prod <- data.hourly %>% 
  filter(variable == "generators_t_p") %>% 
  rename(generators_t_p = value) %>% 
  select(!c("unit", "variable"))

data.temp <- full_join(data.cc, data.prod) %>% 
  group_by(region, bus, technology, generators_p_nom_opt, generators_capital_cost) %>% 
  summarise(gen = sum(generators_t_p)) %>% 
  mutate(value = generators_capital_cost * generators_p_nom_opt / gen,
         variable = "capital cost") %>% 
  ungroup() %>% 
  select(region, bus, technology, variable, value)

# Marginal costs
data.mc <- data.yearly %>% 
  filter(variable == "generators_marginal_cost") %>% 
  mutate(variable = "marginal cost") %>%
  select(!unit)

data.plot.lcoe <- bind_rows(data.mc, data.temp)

source("position_stack_and_nudge.R")

p <- ggplot() +
  geom_bar(data = data.plot.lcoe,
           mapping = aes(x = bus, y = value, fill = variable),
           stat = "identity",
           width = 0.35,
           position = position_stack_and_nudge(x = -0.2)) +
  scale_fill_discrete(name = "LCOE components") +
  new_scale_fill() + 
  geom_bar(data = data.mv,
           mapping = aes(x = bus, y = value, fill = variable),
           stat = "identity",
           width = 0.35,
           position = position_stack_and_nudge(x = 0.2)) +
  scale_fill_manual(values = c("market value" = "gray50")) +
  facet_wrap(~technology)

print(p)

```

## VRE value factors

```{r}

data.plot.mv <- data.mv %>% 
  filter(technology %in% c("offwind-ac", "offwind-dc", "onwind", "solar")) %>% 
  select(region, bus, technology, value)

data.plot.price <- data.hourly %>% 
  filter(variable == "buses_t_marginal_price") %>% 
  rename(price = value) %>% 
  select(region, bus, snapshots, price)

data.plot.load <- data.hourly %>% 
  filter(variable == "loads_t_p") %>% 
  rename(load = value) %>% 
  select(region, bus, snapshots, load)

data.plot.price.weighted <- full_join(data.plot.load, data.plot.price) %>% 
  group_by(region, bus) %>% 
  summarise(price.weighted = sum(load*price)/sum(load))

data.plot <- full_join(data.plot.mv, data.plot.price.weighted) %>% 
  mutate(value.factor = value / price.weighted)

p <- ggplot() + 
  geom_bar(data = data.plot,
           mapping = aes(x = bus, y = value.factor, fill = technology),
           stat = "identity",
           position = "dodge")

print(p)

```


## LCOE and prices

Without investment, LCOE = short-term marginal costs (O&M, fuel, CO2)

```{r}

# Full-year electricity price (across buses and time, weighted by power)
# This is the marginal price (output!)
data.plot.price.year <- data.plot.price.year

# Full-year marginal costs (across technologies and time, weighted by power)
# This is the marginal cost (input!)
data.mc <- data.yearly %>% 
  filter(variable == "generators_marginal_cost") %>% 
  group_by(region, bus, technology) %>% 
  transmute(marginal_cost = value)
  
data.power <- data.hourly %>% 
  filter(variable == "generators_t_p") %>% 
  group_by(region, bus, technology, snapshots) %>% 
  transmute(power = value)

data.plot.mc <- full_join(data.mc, data.power) %>% 
  group_by(region) %>% 
  summarise(mc.weighted = sum(marginal_cost*power)/sum(power))

data.plot <- bind_rows(
  data.plot.price.year %>%
    rename(value = price.weighted) %>% 
    mutate(var = "electricity price (weighted)"),
  data.plot.mc %>% 
    rename(value = mc.weighted) %>% 
    mutate(var = "marginal cost (weighted)")
)

p <- ggplot(data = data.plot) + 
  geom_bar(aes(x = region, y = value, fill = var),
           stat = "identity",
           position = "dodge")

print(p)

```



# Save output as GDX

## Using gdx

```{r}

out <- c(0.5, 0.3)

gdxdata <- list()
gdxdata$name <- "pm_pypsa_capfac"
gdxdata$type <- "parameter"
gdxdata$varTypeText <- "TEST"
gdxdata$domains <- c("all_regi")

attributes(out) <- c(attributes(out), list(gdxdata = gdxdata))

writeGDX(out, file="test.gdx", period_with_y = FALSE)

```

## Using gdxdt

```{r}

dt <- data.table(
	col1=c("a", "b", "c"),
	col2=c("d", "e", "f"))
dt[, value:=runif(1), by="col1"]

writegdx.parameter("test.gdx", dt,
	name="test",
	valcol = "value",
	uelcols = c("col1", "col2"))

```


